---
title: "ds202_hw3"
author: "Danny Nguyen"
date: "February 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

```{r}
data(tips, package="reshape2")
summary(tips)
print(tips)
```

1. How many parties did the waiter serve? Store the number in numParty and print.
If we are referring parties to the number of observations then the number of parties the waiter served was 244.
```{r}
numParty <- 244
print(numParty)
```
2. What are the types of variables in the data frame tips?
The types of the variables in the data frame tips are double and integer.
```{r}
typeof(tips$total_bill)
typeof(tips$tip)
typeof(tips$sex)
typeof(tips$smoker)
typeof(tips$day)
typeof(tips$time)
typeof(tips$size)
```
3. Create a vector named day containing the day of visits in tips. The factor levels should be ordered from Thursday to Sunday. Also print the variable.
```{r}
day <- tips$day
print(day)
```
4. Create a data frame named female5 containing the meal paid by a female payer in a party with size greater than or euqal to 5. Also print the data frame.
```{r}
female5 <- data.frame(tips$sex == "Female", tips$size >= 5)
print(female5)
```
5. How does the tipping amount (tip) depend on the overall bill (total_bill)? Describe the relationship in words. Describe at least two types of anomalies in the plot. What do they mean? Use the ggplot2 package to make a chart.
Based on the chart, it seems there is a possible positive correlation between the total_bill and tipping amount. One anomaly is that there seems to be a straight linear trend around the 2 and 3 dollar marks this could be because maybe those are the average tipping amounts. The other anomaly I could tell would be the density of the points, it is a lot more clustered at the lower amounts but more spread out at the higher amounts.
```{r}
ggplot(data = tips, aes(x = total_bill, y = tip)) + geom_point()
```
6. Introduce a variable tiprate into the data set, that incorporates the rate of tips. What is the average rate for tips?
```{r}
tiprate <- mean(tips$tip)
print(tiprate)
```
7. Make a boxplot of the tip rate by time. The x-axis should be ordered by lunch and then dinner. Use the ggplot2 package to make a chart.
```{r}
ggplot(data = tips, aes(x = tiprate, y = time)) + geom_boxplot() + coord_flip()
```
8. How does smoking behavior and gender of the person who pays impact the relationship between tip and total bill? Find a visualization that incorporates all four variables. Interpret the result. 
Well males and females that don't smoke seem to be the more frequent tippers, but those that do smoke are more spread out in terms of tipping and total bill.
```{r}
ggplot(data = tips, aes(x=total_bill, y=tip, color=as.factor(sex))) +
  geom_point() + 
  facet_grid(. ~ smoker)
```
9. Use ggplot2 to find a graphical summary of the relationship between day of the week and gender of the person paying the bill. What can you say about this relationship?
Well other than Saturday, it seems that there are more males paying the bill than females.
```{r}
ggplot(data = tips, aes(x=day, y=total_bill, color=as.factor(sex))) +
  geom_boxplot()
```